Comparison of dimensionality reduction schemes for parallel global optimization algorithms \thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150}

### 2018-04-08 2018/RSD/MultiEvolvents/RSD.tex
RSD: drop the old introduction. We need a more suitable one

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{cmap}
%%%%%
\usepackage{eqnarray,amsmath}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

@@

Global (or multiextremal) optimization problems are among the most complex problems
in both theory and practice of optimal decision making. In these kinds of problems,
the optimization criterion can have several local optima within the search domain.
The existence of several local optima makes
finding the global optimum difficult essentially, since it requires examining the whole feasible
search domain. The volume of computations for solving global optimization problems
can increase exponentially with increasing number of varied parameters.
\par
These global optimization problem features impose special requirements on the quality
of the optimization methods and on the software to implement these ones. The global
optimization methods should be highly efficient, and the software systems should be developed on
a good professional basis. In general, the global optimization problems can be solved at
a reasonable time by employing parallel computations on modern supercomputing systems only.
\par
The general state of the art in the field of global optimization is presented in a
number of key monographs  \cite{floudasPardGO}, \cite{horstTuyGO}, \cite{locatelliSchoenGO}, \cite{pinterGO}, \cite{strSergGO}, \cite{zilinskTornGO}, \cite{zhigljavskyRandGO}.
The development of optimization methods, which use the high-performance computational
systems to solve the time-consuming global optimization problems, is an area of
intensive research --- see, for instance, \cite{censorZeniosParGO}, \cite{ciegisHentyParGO},
\cite{luqueAlbaGA}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}. The obtained
theoretical results provide the efficient solutions of many applied global
optimization problems in various fields of scientific and technical applications \cite{Famularo1999}, \cite{fasanoPinter2013},
\cite{floudasPardalosGOState}, \cite{Kvasov2015}, \cite{Menniti}, \cite{locatelliSchoenGO},
\cite{luqueAlbaGA}, \cite{pardalosZhigljavskyZilinskas2016}, \cite{pinterGO}.
\par
At the same time, the practical implementation of these global optimization algorithms
within the framework of industrial software systems is quite limited. In many cases,
software implementations are experimental in nature and are used by the developers
themselves to obtain the results from the computational experiments required for the
scientific publications. This situation originates from high development costs of the
professional software systems, which can be used by numerous users. In addition, the global
optimization problems could be solved in an automatic mode rarely because of the
complexity of these ones. The user should actively control the global search
process that implies an adequate level of qualification in the field of optimization
(particularly, the user should know and understand the global optimization methods well).
\par
In this work, the authors consider an approach to minimizing multiextremal functions developed in ...????????
This allows problems to be solved in which function values may not be determined for the entire search domain. Under this approach, solving multidimensional problems is reduced (using Peano-type spacefilling curves) to solving equivalent one-dimensional problems.
It should be noted that standard approaches to algorithm parallelization are not quite applicable to global optimization. For example, the rules for selecting another iteration point are quite simple and do not require parallelization (as overheads associated with organizing parallel computations will nullify any possible acceleration). Some acceleration can be achieved by parallelizing the computa-tion of function values describing the object to be optimized; however, this approach is specific to each individual problem being solved.
The following approach looks more promising. The algorithm can be modified to run several trials in parallel. This approach provides the efficiency (as parallelization is applied to the most computation-intensive part of the problem solving process) and generality (in that it applies to a wide range of global optimiza-tion algorithms). The approach, described in  \cite{Two_Level_Parallel} for unconstrained optimization, was used in this work for parallelizing constrained optimization algorithms.

%------------------------------------------------------------------------------
\section{Statement of Multidimensional Global Optimization Problem}
In this paper, the core class of optimization problems, which can be solved using
Globalizer\cite{globalizerSystem}, is formulated. This class involves the multidimensional global
optimization problems without constraints, which can be defined in the following way:
\begin{equation}
\label{eq:task}
\begin{array}{cr}\\
  \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
  D=\{y\in \mathbf{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
\end{array}
\end{equation}
with the given boundary vectors  $a$ and  $b$. It is supposed, that the objective function \(\varphi(y)\) satisfies the Lipschitz condition
\begin{equation}
\label{eq:lip}
|\varphi(y_1)-\varphi(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
\end{equation}
where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbf{R}^N\) space.
\par
Usually, the minimized function \(\varphi(y)\) is defined as a computational procedure,
according to which the value \(\varphi(y)\) can be calculated for any vector \(y\in D\)
(let us further call such a calculation \textit{a trial}). It is supposed that this procedure
is a time-consuming one. As a result, the overall time of solving the optimization
problem (\ref{eq:task}) is determined, first of all by the number of executed trials.
It should also be noted that the requirement of the Lipschitz condition (\ref{eq:lip})
is highly important, since an estimate of the global minimum can be constructed on the
basis of a finite number of computed values of the optimized function only in this case .
\par
As it has been shown earlier by many researchers
(see, for instance, \cite{floudasPardalosGOState}, \cite{horstTuyGO}, \cite{pinterGO}, \cite{strSergGO}),
finding the numerical estimate of the global optimum implies constructing a coverage of
the search domain \(D\). As a result, the computational costs of solving the global
optimization problems are readily very high even for a small number of varied parameters
(the dimensionality of the problem). A notable reduction in the volume of computations
can be achieved when the coverage of the search domain is non-uniform, i. e. the series
of trial points is only dense in a vicinity of the global optimum point. The construction
of such a non-uniform coverage could be provided in an adaptive way, when the selection
of the next trial points is determined by using the search information (the preceding
trial points and the values of the minimized function at these points) obtained in the course
of computations. This necessary condition complicates considerably the computational schemes
of global optimization methods since it implies a complex analysis of a large amount
of multidimensional search information. As a result, many optimization algorithms use
various approaches to the dimensional reduction \cite{pinterGO}, \cite{sergeyevStronginLera2013}, \cite{strongin1978}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.

@@

Within the framework of the information-statistical global optimization theory,
the Peano space-filling curves (or evolvents) \(y(x)\) mapping the interval \([0,1]\)
onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
reduction \cite{sergeyevStronginLera2013}, \cite{strongin1978}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
\par
As a result of the reduction, the initial multidimensional global optimization
problem (\ref{eq:task}) is reduced to the following one-dimensional problem:
\begin{equation}
\label{eq:oneDimTask}
\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}.
\end{equation}
\par
It is important to note that this dimensionality reduction scheme transforms the minimized
Lipschitzian function from (\ref{eq:task}) to the corresponding one-dimensional
function \(\varphi(y(x))\), which satisfies the uniform H{\"o}lder condition, i. e.
\begin{equation}
\label{eq:holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
\end{equation}
where the constant $H$ is defined by the relation \(H=2L\sqrt{N+3}\), \(L\) is the Lipschitz
constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem (\ref{eq:task}).
\par
The algorithms for the numerical construction of the Peano curve approximations are
given in \cite{strSergGO}. As an illustration, an approximation of the Peano curve
for the third density level is shown in Figure \ref{fig:peanoC}. Figure \ref{fig:peanoC}
demonstrates the movement order in a two-dimensional domain to construct the Peano
curve approximation; the precision of the Peano curve approximation is determined by the
density level used in the construction.
\begin{figure}
    \centering
    \includegraphics[width=0.85\textwidth]{pictures/peanoC.eps}
    \caption{A Peano curve approximation for the third density level}
    \label{fig:peanoC}
\end{figure}

\par
The computational scheme obtained as a result of the dimensionality reduction consists of the following:
%(see Figure \ref{fig:peanoCUsage}):
\begin{itemize}
  \item The optimization algorithm performs the minimization of the reduced one-dimensional
  function \(\varphi(y(x))\) from (\ref{eq:oneDimTask}),
  \item After determining the next trial point \(x\), a multidimensional image \(y\) is calculated by using the
mapping \(y(x)\),
  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the point \(y\in D\),
  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
\end{itemize}

%\begin{figure}
%    \centering
%    \includegraphics[width=0.55\textwidth]{pictures/peanoCUsage.eps}
%    \caption{The computational scheme for obtaining the value of the reduced one-dimensional function \(\varphi(y(x))\)}
%    \label{fig:peanoCUsage}
%\end{figure}

@@

Let us consider a family of the hypercubes
\begin{equation}\label{6_hypercubes}
D_l= \left\{y \in R^N: -2^{-1} \leq y_i+2^{-l} \leq 3 \cdot 2^{-1},\ 1\leq i\leq N\right\},\ 0 \leq l \leq L,
\end{equation}
where the hypercube $D_{l+1}$ is obtained by the translation of the hypercube $D_l$ along the main diagonal with displacement equal to $2^{-l}$ in each coordinate. In Fig.~\ref{fig:shifted_ev} the hypercubes $D_0,...,D_3$ for the case $N=2,\ L=3$ are presented.

@@

Let us assume that the evolvent $y^0(x)$  maps the interval $[0,1]$ onto the hypercube $D_0$ from (\ref{6_hypercubes}), i.e.,
\[
D_0 = \left\{y^0(x) : x \in [0,1]\right\}.
\]
Then, the evolvents $y^l(x)=\left\{y_1^l(x),...,y_N^l(x)\right\}$, the coordinates of which are defined by the conditions
\[
y_i^l(x)=y_i^{l-1}(x)+2^{-l},\ 1\leq i\leq N, \ 1\leq l\leq L,
\]
map the interval $[0,1]$ onto the corresponding hypercubes $D_l,\ 1\leq l \leq L$. In Fig.~\ref{fig:shifted_ev} the image of the interval $[0,1]$ obtained by the curve $y^0(x),\ x\in [0,1],$ is shown as the dashed line. Since the hypercube $D$ from (\ref{eq:task}) is included in the common part of the family of hypercubes (\ref{6_hypercubes}) (the boundaries of hypercube $D$ are highlighted in Fig.~\ref{fig:shifted_cube}), having introduced an additional constraint function
\begin{equation}\label{6_g0}
g_0(y)=\max\left\{\left|y_i\right| - 2^{-1}:\ 1\leq i\leq N\right\},
\end{equation}
one can present the initial hypercube $D$ in the form
\[
D=\left\{y^l(x):\; x\in [0,1],\ g_0(y^l(x))\leq 0 \right\},\ 0\leq l \leq L,
\]
i.e., $g_0(y) \leq 0$ if $y\in D$ and $g_0(y)>0$ otherwise. Consequently, any point $y \in D$ has its own preimage $x^l \in [0,1]$ for each mapping $y^l(x),\ 0\leq l\leq L$.

The problems (\ref{6_problem_l}) correspond to the domains $Q_0^l=[0,1]$ and $Q_{j+1}^l, 0 \leq j \leq m,$ defined by the expression
\[
Q_{j+1}^l = \left\{x \in Q_j^l:g_j(y^l(x))\leq 0\right\},\ 0\leq j\leq m,
\]
for the corresponding evolvents $y^l(x)$. The application of a multiple mapping defines the following relation for the nearness in the multidimensional search domain and in the one-dimensional one.

\begin{theorem}
Let a point $y^\ast$ from the domain $D$ be contained in the line segment with the end-points $y',y'' \in D$ meeting the requirements
\[
\left|y'_j - y''_j\right|\leq 2^{-p},\ y'_i=y''_i=y^\ast_i, \ 1\leq i \leq N,\ i \neq j,
\]
where $p$ is an integer  and $1\leq p \leq L$, i.e., the line segment is collinear with the $j$-th axis in $R^N$. Then, there exist at least one mapping $y^l(x),\ 0\leq l\leq L$, and the preimages $x^\ast,\; x',\; x''\in [0,1]$ such that
\[
y^\ast = y^l(x^\ast),\ y'=y^l(x'),\ y'' = y^l(x'')
\]
and
\[
\max \left\{ \left|x'-x^\ast\right|,\; \left|x''-x^\ast\right|,\; \left|x'-x''\right|\right\} \leq 2^{-pN}.
\]
\end{theorem}
\begin{proof}
Proof of this theorem is given in \cite{strSergGO}.
\qed
\end{proof}

The conditions of the theorem single out a specific vicinity of the point $y^\ast$. This vicinity includes only the points, which can be obtained by the shift of $y^\ast$ parallel to one of the coordinate axes with a displacement not more than $2^{-p}$. By changing  $j,\ 1\leq j\leq N,$ in the theorem conditions it is possible to obtain the neighbours in any $N$ coordinate directions. According to the statement, the closeness of the points in the $N$-dimensional space in a particular direction will be reflected by the closeness of their preimages in one of the univariate problems. The information on the closeness of the points results, first, in more precise estimate of Lipschitz constants and, second, in the increase of the characteristics of the intervals, the images of the end points of which are close to each other in the $N$-dimensional space.

%------------------------------------------------------------------------------
\subsection{Rotated evolvents}
The application of the scheme for building the multiple evolvents (hereinafter called the shifted evolvents or $S$-evolvents) described in Subsection \ref{sec:shifted} allows to preserve the information on the nearness of the points in the multidimensional space and, therefore, to provide more precise (as compared to a single evolvent) estimate of Lipschitz constant in the search process. However, this approach has serious restrictions, which narrow the applicability of the parallel algorithms, designed on the base of the $S$-evolvents.

Because a shifted evolvent is built as a result of the shift of a hypercube $D$ along the main diagonal, and the shift step decreases 2 times for the building of each subsequent mapping, the number of such shifts is limited by the evolvent density $m$. In the case when the shift step is less than $2^{-m}$, the next evolvent coincides with the previous one. Thus, the applicable number of the evolvents $L$ and, hence, the number of processors are limited by $m$ ($L\leq m$), where $m$ is the density of the evolvents.

Also, it follows from the algorithm for building the shifted evolvents that solving the initial problem in the domain $D$ is reduced to solving a set of problems in the extended domains $D_l$ from (\ref{6_hypercubes}) with additional constraint (\ref{6_g0}). As a result, the structure of the search domain becomes more complicated, and the exponential decrease of the volume of the search domain $D$ in relation to the volumes of the domains $D_l$ with increasing dimensionality of the problem takes place.

To overcome the disadvantages mentioned above and to preserve the information on the nearness of the points in the $N$-dimensional space, a novel scheme of building of the multiple mappings is proposed. The building of a set of Peano curves not by the shift along the main diagonal of the hypercube but by rotation of the evolvents around the coordinate origin is a distinctive feature of the proposed scheme \cite{Gergel2009}. In the initial non-rotated mapping for close points $y', y''$ in the multidimensional space their preimages  $x', x''$ in the interval $[0,1]$ can be far away from each other. In the rotated scheme there exists a mapping $y^i(x)$ according to which preimages $x', x''$ will be located nearer. The evolvents generated according to the novel scheme will be hereinafter called the rotated evolvents or $R$-evolvents.

In Fig.~\ref{6_fig_9} two evolvents being the approximations to Peano curves for the case $N=2$ are presented as an illustration.

@@

Taking into account the initial mapping, one can conclude that current implementation of the method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals. Moreover, the additional constraint  $g_0(y) \leq 0$ with $g_0(y)$ from (\ref{6_g0}), which arises in shifted evolvents, is absent. This method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.

The use of the set of mappings $Y_L(x)=\{y^1(x),...,y^L(x)\}$ results in the appearing of the corresponding set of the one-dimensional multiextremal problems
\begin{equation}\label{6_problem_lr}
  \min{\left\{\varphi(y^l(x)):x\in [0,1], \; g_j(y^l(x))\leq 0, \; 1 \leq j \leq m\right\}}, \ 1 \leq l \leq L.
\end{equation}
Each problem from this set can be solved independently. Any computation of the value $z=g_\nu(y'),\ y'=y^i(x')$ of the function $g_\nu(y)$ in the $i$-th problem can be interpreted as a computation of the value $z=g_\nu(y'),\ y'=y^s(x'')$ for any other $s$-th problem without time-comsuming computations of the functions $g_\nu(y)$. Such information integrity allows to solve  initial problem (\ref{eq:task}) by solving $L$ problems (\ref{6_problem_lr}) on a set of intervals $[0,1]$ in parallel by the index method.

@@

%\begin{figure}[t]
%  \centering
%  \includegraphics[width=0.6\linewidth]{pictures/noninjective.pdf}
%  \caption{Numerical non injective evolvent}
%  \label{fig:noninjective}
%\end{figure}

@@

%\begin{figure}[t]
%  \centering
%  \includegraphics[width=0.6\linewidth]{pictures/smooth.pdf}
%  \caption{Smooth evolvent}
%  \label{fig:smooth}
%\end{figure}

@@

The information-statistical theory of global optimization formulated in \cite{strongin1978}, \cite{strSergGO} has
served as a basis for the development of a large number of efficient multiextremal optimization
methods \cite{barkalovGergel2014}, \cite{gergel1996}, \cite{gergel1997}, \cite{grishaginStrongin1984}, \cite{Pizzuti}, \cite{sergeyev1995}, \cite{sergeyev1999}, \cite{sergeyevGrishagin2001}, \cite{sergeyevStronginLera2013}, \cite{Famularo2001}.
\par
The optimization methods applied in Globalizer are based on the MAGS method, which can be presented as follows --- see \cite{strongin1978}, \cite{strSergGO}.
\par
Let us introduce a simpler notation for the optimization problem being solved
\begin{equation}
\label{eq:oneDimFunc}
f(x) = \varphi(y(x)):x\in [0,1].
\end{equation}
\par
The initial iteration of the algorithm is performed at an arbitrary point \mbox{\(x^1\in(0,1)\)}.
Then, let us suppose that \(k\), \(k\ge 1\), optimization iterations have been completed already.
The selection of the trial point \(x^{k+1}\) for the next iteration is performed according to the following rules.
\par
\textit{Rule 1}. Renumber the points of the preceding trials by the lower indices in order of increasing value of coordinates
\begin{equation}
  \label{step1}
0=x_0<x_1<...<x_{k+1}=1,
\end{equation}
the points \(x_0\), \(x_{k+1}\) were introduced additionally for the convenience of further
explanation, the values of the minimized function \(z_0\), \(z_{k+1}\) at these points are undefined.
\par
\textit{Rule 2}. Compute the current estimate of the H{\"o}lder constant \(H\) from (\ref{eq:holder})
\begin{equation} \label{step2}
M=\max_{1 < i\leq k}\frac{|z_i-z_{i-1}|}{\rho_i}, \;
m = \left\{
   \begin{array}{lr}
     rM, & M > 0,\\
     1, & M = 0,
   \end{array}
  \right.
\end{equation}
as the maximum of the relative differences of the minimized function values on the
set of previously executed trial points \(x_i,1\leq i\leq k\) from (\ref{step1}).
Hereafter, \(\rho_i=(x_i-x_{i-1})^\frac{1}{N},1\leq i\leq k+1\). The
constant \(r\), \(r>1\), is the reliability parameter of the algorithm.
\par
\textit{Rule 3}. Compute the characteristics \(R(i)\) for each interval \((x_{i-1},x_i),1\leq i\leq k+1\), where
\[
R(i)=2\rho_i-4\frac{z_i}{m},\quad i=1,
\]
\begin{equation} \label{step3}
R(i)=\rho_i+\frac{(z_i-z_{i-1})^2}{m^2\rho_i}-2\frac{z_i+z_{i-1}}{m},\quad 1<i<k+1, \\
\end{equation}
\[
R(i)=2\rho_{i}-4\frac{z_{i-1}}{m},\quad i=k+1.
\]

\par
\textit{Rule 4}. Determine the interval with the maximum characteristic
\begin{equation} \label{step4}
R(t)=\max_{1\leq i \leq k+1}R(i).
\end{equation}
\par
\textit{Rule 5}. Execute a new trial at the point \(x^{k+1}\) located within the interval
with the maximum characteristic from (\ref{step4})
\begin{equation} \label{step5}
  x^{k+1}=\frac{x_t+x_{t-1}}{2}-\mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{r|z_{t}-z_{t-1}|}{m}\right]^N,\; \textrm{ if } 1<t<k+1,
\end{equation}
\[
  x^{k+1}=\frac{x_t+x_{t-1}}{2},\; \textrm{ if } t=1 \textrm{ or } t=k+1.
\]

\par
The stopping condition, which terminated the trials, is defined by the inequality
\begin{equation}
  \label{eq:stop_1}
\rho_t<\varepsilon
\end{equation}
for the interval with the maximum characteristic from (\ref{step4}) and \(\varepsilon >0\) is the predefined
accuracy of the optimization problem solution. If the stopping condition is not satisfied,
the index \(k\) is incremented by 1, and the new global optimization iteration is executed.
\par
In order to explain the algorithm presented above, let us note the following.
The characteristics \(R(i), 1\leq i\leq k+1\), calculated according to (\ref{step3}) could
be interpreted as some measures of importance of the intervals with respect to the
expected location of the global minimum point. Thus, the rules (\ref{step4}) and (\ref{step5}) for selecting
the interval of the next trial become more clear --- the point of every next
global optimization iteration is selected within the interval, where the global minimum
point can be found most likely.
\par
The convergence conditions of the described algorithm are given, for example, in \cite{strSergGO}.

%------------------------------------------------------------------------------
\subsection{Параллельные множественные отображения}
\label{sec:parallel_evolvents}
Using the multiple mapping allows solving initial problem (1) by parallel solving the problems
\[
\min\{\varphi(y^s(x)):x\in [0,1]\}, 1≤s≤S
\]
on a set of intervals [0,1] by the index method. Each one-dimensional problem is solved on a separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by particular processor are interpreted as the results of the trials in the rest problems (in the corresponding points \(x^(k_1 ),…,x^(k_S ))\). In this approach, a trial at the point \(x^k \in [0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of operations.
\par
1. Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
\par
2. Inform the rest of processors about the start of the trial execution at the point\( y^k\) (the blocking of the point \(y^k\) ).
\par
3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1≤s≤S\), of the point \(y^k\) and interpret the trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points \(x{}^{k_1} ,…,x{}^{k_s} \)
\par
4. Inform the rest of processors about the trial results at the point \(y^k\).
\par

@@

\bibitem{barkalovGergel2014}
\newblock K. Barkalov and V. Gergel,
\newblock \emph{\emph{Multilevel scheme of dimensionality reduction for parallel global search algorithms}},
\newblock in \emph{Proceedings of the 1st International Conference on Engineering and Applied Sciences Optimization}, (2014), 2111--2124.

\bibitem{barkalovGergel2015}% (MR3539505) [10.1007/s10898-016-0411-y]
\newblock K. Barkalov and V. Gergel,
\newblock \emph{\emph{Parallel global optimization on GPU}},
\newblock \emph{J. Glob. Optim.}, \textbf{66} (2016), 3--20.

\bibitem{barkalovGergelLebedev2015} %[10.1007/978-3-319-21909-7_31]
\newblock K. Barkalov, V. Gergel and I. Lebedev,
\newblock \emph{\emph{Use of Xeon Phi coprocessor for solving global optimization problems}},
\newblock \emph{LNCS}, \textbf{9251} (2015), 307--318.

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
%%%%%ДОБАВИЛ ДЛЯ РУССКОГО ТЕКСТА
\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{cmap}
%%%%%
\usepackage{eqnarray,amsmath}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\setlength{\textfloatsep}{4pt plus 1.0pt minus 1.5pt}
\setlength{\floatsep}{4pt plus 1.0pt minus 1.5pt}
%\setlength{\abovedisplayskip}{3pt}
%\setlength{\belowdisplayskip}{3pt}

@@

%------------------------------------------------------------------------------
\section{Statement of Multidimensional Global Optimization Problem}
In this paper, the core class of optimization problems, which can be solved using
Globalizer\cite{globalizerSystem}, is formulated. This class involves the multidimensional global
optimization problems without constraints, which can be defined in the following way:
\begin{equation}
\label{eq:task}
\begin{array}{cr}\\
  \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
  D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
\end{array}
\end{equation}
with the given boundary vectors  $a$ and  $b$. It is supposed, that the objective function \(\varphi(y)\) satisfies the Lipschitz condition
\begin{equation}
\label{eq:lip}
|\varphi(y_1)-\varphi(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
\end{equation}
where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbf{R}^N\) space.
\par
Usually, the minimized function \(\varphi(y)\) is defined as a computational procedure,
according to which the value \(\varphi(y)\) can be calculated for any vector \(y\in D\)
(let us further call such a calculation \textit{a trial}). It is supposed that this procedure
is a time-consuming one. As a result, the overall time of solving the optimization
problem (\ref{eq:task}) is determined, first of all by the number of executed trials.
It should also be noted that the requirement of the Lipschitz condition (\ref{eq:lip})
is highly important, since an estimate of the global minimum can be constructed on the
basis of a finite number of computed values of the optimized function only in this case .

@@

Within the framework of the information-statistical global optimization theory,
the Peano space-filling curves (or evolvents) \(y(x)\) mapping the interval \([0,1]\)
onto an \(N\)-dimensional hypercube \(D\) unambiguously are used for the dimensionality
reduction \cite{sergeyevStronginLera2013}, \cite{strongin1978}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
\par
As a result of the reduction, the initial multidimensional global optimization
problem (\ref{eq:task}) is reduced to the following one-dimensional problem:
\begin{equation}
\label{eq:oneDimTask}
\varphi(y(x^*))=\min\{\varphi(y(x)):x\in [0,1]\}.
\end{equation}
\par
It is important to note that this dimensionality reduction scheme transforms the minimized
Lipschitzian function from (\ref{eq:task}) to the corresponding one-dimensional
function \(\varphi(y(x))\), which satisfies the uniform H{\"o}lder condition, i. e.
\begin{equation}
\label{eq:holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
\end{equation}
where the constant $H$ is defined by the relation \(H=2L\sqrt{N+3}\), \(L\) is the Lipschitz
constant from (\ref{eq:lip}), and \(N\) is the dimensionality of the optimization problem (\ref{eq:task}).
\par
The algorithms for the numerical construction of the Peano curve approximations are
given in \cite{strSergGO}.

\par
The computational scheme obtained as a result of the dimensionality reduction consists of the following:
\begin{itemize}
  \item The optimization algorithm performs the minimization of the reduced one-dimensional
  function \(\varphi(y(x))\) from (\ref{eq:oneDimTask}),
  \item After determining the next trial point \(x\), a multidimensional image \(y\) is calculated by using the
mapping \(y(x)\),
  \item The value of the initial multidimensional function \(\varphi(y)\) is calculated at the point \(y\in D\),
  \item The calculated value \(z=\varphi(y)\) is used further as the value of the reduced one-dimensional function \(\varphi(y(x))\) at the point \(x\).
\end{itemize}

@@

Such set of evolvents can be produced by shifting the source evolvent $y^0(x)$ by $2^{-l},0 \leq l \leq L$ on each coordinate. Each evolvent has it's own corresponding hypercude $D_l= \left\{y \in R^N: -2^{-1} \leq y_i+2^{-l} \leq 3 \cdot 2^{-1},\ 1\leq i\leq N\right\},\ 0 \leq l \leq L$.

In Fig.~\ref{fig:shifted_ev} the image of the interval $[0,1]$ obtained by the curve $y^0(x),\ x\in [0,1],$ is shown as the dashed line. Since the hypercube $D$ from (\ref{eq:task}) is included in the common part of the family of hypercubes $D_l$ (the boundaries of hypercube $D$ are highlighted in Fig.~\ref{fig:shifted_cube}), having introduced an additional constraint function
\begin{equation}\label{6_g0}
g_0(y)=\max\left\{\left|y_i\right| - 2^{-1}:\ 1\leq i\leq N\right\},
\end{equation}
one can present the initial hypercube $D$ in the form
\[
D=\left\{y^l(x):\; x\in [0,1],\ g_0(y^l(x))\leq 0 \right\},\ 0\leq l \leq L,
\]
i.e., $g_0(y) \leq 0$ if $y\in D$ and $g_0(y)>0$ otherwise. Consequently, any point $y \in D$ has its own preimage $x^l \in [0,1]$ for each mapping $y^l(x),\ 0\leq l\leq L$.

@@

%------------------------------------------------------------------------------
\subsection{Rotated evolvents}
The application of the scheme for building the multiple evolvents (hereinafter called the shifted evolvents or $S$-evolvents) described in Subsection \ref{sec:shifted} allows to preserve the information on the nearness of the points in the multidimensional space and, therefore, to provide more precise (as compared to a single evolvent) estimate of Lipschitz constant in the search process. However, this approach has serious restrictions, which narrow the applicability of the parallel algorithms, designed on the base of the $S$-evolvents (see te end of the section \ref{sec:seq_comp}).

To overcome the $S$-evolvent and to preserve the information on the nearness of the points in the $N$-dimensional space, a novel scheme of building of the multiple mappings is proposed. The building of a set of Peano curves not by the shift along the main diagonal of the hypercube but by rotation of the evolvents around the coordinate origin is a distinctive feature of the proposed scheme \cite{Gergel2009}. In the initial non-rotated mapping for close points $y', y''$ in the multidimensional space their preimages  $x', x''$ in the interval $[0,1]$ can be far away from each other. In the rotated scheme there exists a mapping $y^i(x)$ according to which preimages $x', x''$ will be located nearer. The evolvents generated according to this scheme are called the rotated evolvents or $R$-evolvents.

In Fig.~\ref{6_fig_9} two evolvents being the approximations to Peano curves for the case $N=2$ are presented as an illustration.
Taking into account the initial mapping, one can conclude that current implementation of the method allows to build up to $N(N-1)+1$ evolvents for mapping the $N$-dimensional domain onto the corresponding one-dimensional intervals. Moreover, the additional constraint  $g_0(y) \leq 0$ with $g_0(y)$ from (\ref{6_g0}), which arises in shifted evolvents, is absent. This method for building a set of mappings can be ``scaled'' easily to obtain more evolvents (up to $2^N$) if necessary.

@@

@@

@@

@@

The optimization methods applied in Globalizer to solve the reduced problem (\ref{eq:oneDimTask}) are based on the MAGS method, which can be presented as follows --- see \cite{strongin1978}, \cite{strSergGO}.
\par
The initial iteration of the algorithm is performed at an arbitrary point \mbox{\(x^1\in(0,1)\)}.
Then, let us suppose that \(k\), \(k\ge 1\), optimization iterations have been completed already.
The selection of the trial point \(x^{k+1}\) for the next iteration is performed according to the following rules.

\textit{Rule 1}. Renumber the points of the preceding trials by the lower indices in order of increasing value of coordinates
$0=x_0<x_1<...<x_{k+1}=1$.

\textit{Rule 2}. Compute the characteristics \(R(i)\) for each interval \((x_{i-1},x_i),1\leq i\leq k+1\).

\textit{Rule 4}. Determine the interval with the maximum characteristic $R(t)=\max_{1\leq i \leq k+1}R(i)$.

\textit{Rule 5}. Execute a new trial at the point \(x^{k+1}\) located within the interval with the maximum characteristic from the previous step
  $x^{k+1}=d(x_t)$.

The stopping condition, which terminated the trials, is defined by the inequality $\rho_t<\varepsilon$
for the interval with the maximum characteristic from Step 4 and \(\varepsilon >0\) is the predefined
accuracy of the optimization problem solution. If the stopping condition is not satisfied,
the index \(k\) is incremented by 1, and the new global optimization iteration is executed.

The convergence conditions and exact formulas for descision rules $R(i)$ and $d(x)$ of the described algorithm are given, for example, in \cite{strSergGO}.

%------------------------------------------------------------------------------
\subsection{Parallel algorithm exploiting a set of evolvents}
\label{sec:parallel_evolvents}
Using the multiple mapping allows solving initial problem (\ref{eq:task}) by parallel solving the problems
\[
\min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leqslant s\leqslant S
\]
on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by particular processor are interpreted as the results of the trials in the rest problems (in the corresponding points \(x^(k_1 ),\dots,x^(k_S))\). In this approach, a trial at the point \(x^k \in [0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of operations.
\par
1. Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
\par
2. Inform the rest of processors about the start of the trial execution at the point\( y^k\) (the blocking of the point \(y^k\) ).
\par
3. Determine the preimages \(x{}^{k_s}  \in [0,1], 1≤s≤S\), of the point \(y^k\) and interpret the trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points \(x{}^{k_1} ,…,x{}^{k_s} \)
\par
4. Inform the rest of processors about the trial results at the point \(y^k\).
\par

@@
